---
title: "hw05"
author: "Yeeun Lee"
output: html_document
---

# Exercise 1: Explain the value of the here::here package 
"here::here()" provides a simple way to detect the root directory as well as set a working directory. What I like about "here" package most is that I can easily detect root directory of a file or data or work with paths, and those do not need to correspond to the working directory. That is the big difference between 'hear' and 'setwd' or 'getwd'. 
Moreover, by setting relative paths, users do not need to specify all paths explicitly, which makes R project accessible in a platform-independent way.


# Exercise 2: Factor management

Step1. Selecting datase: gapminder
```{r, include=FALSE, message=FALSE}
library(gapminder)
library(tidyverse)
library(dplyr)
library(forcats)
library(ggplot2)
library(here)
library(plotly)
library(tidyverse)
```

Step 2: Check levels
I will first check (1) whether continent is factor or character and (2) what are the levels
```{r}
class(gapminder$continent)
levels(gapminder$continent)
```

Step 3: Drop "Oceania""
```{r, warning = FALSE, message = FALSE}
gap_drop_oceania <- gapminder %>% 
  filter(!continent == "Oceania")
gap_drop_oceania  %>% 
  droplevels()
gap_drop_oceania

```

Step 4: Reorder the levels of country or continent. 
I will select 4 East Asia countries and reorder the levels of country by gdpPercap.  
```{r}
eastasia <- c("Japan", "Korea, Rep.", "China", "Korea, Dem. Rep.")
east_asia <- gapminder %>%
  filter(country %in% eastasia)

ggplot(east_asia, aes(x=year, y= gdpPercap, 
            color = fct_reorder2(country, year, gdpPercap))) +
  geom_line()+ 
  labs (color="country")

```


# Exercise 3: File input/output (I/O) 


Step 1. Export the East_asian file

```{r}
write_csv(east_asia, here("east_asia.csv"))
```

Step 2. Import the file again.

```{r, message=FALSE, warning= FALSE, results=FALSE}
read_csv(here::here("east_asia.csv"))
```


# Exercise 4: Visualization design

Step 1: Selecting one previous plot: (hw 03)How is life expectancy changing over time on different continents?
```{r}
plot1 <- gapminder %>%
  group_by(continent, year) %>%    
  mutate(life_expectancy = mean(lifeExp)) %>%
  ggplot(aes(year, life_expectancy, colour = continent)) + geom_line(alpha = 0.4)

plot1
```

Step 2: Revising the plot: chaging the labels; adding the title; making it interactive using plotly package 
```{r}
plot2 <- gapminder %>%
  group_by(continent, year) %>%    
  mutate(life_expectancy = mean(lifeExp)) %>%
  ggplot(aes(year, life_expectancy, colour = continent)) +
  geom_line(alpha = 0.4) +
  labs(title="Life expectancy changes by continent", x="Year", y="Life expectancy") +
  theme(panel.grid.minor = element_blank())

plot2 %>%
  ggplotly()

```

Step 3: Placing these plots side-by-side:

```{r}
(plot_comparison<- gridExtra::grid.arrange(plot1, plot2, ncol = 2))
```

# Exercise 5: Writing figures to file

```{r}
ggsave("life_expectancy.png", plot = plot_comparison, width = 20, height = 10, dpi = 300)
```


